Initial Analysis

### 2019-05-28 Initial_Analysis.Rmd
Fixed coordinates issues, ready to run!

GetSppDist=function(Spp,Model){

  Distpath <- paste(Data_Path,"Distribution_Data/",sep="")
  INDEX <- seq(1,259200,1)

@@

  # Occurence
  if(Model == "Occurence"){
    File_Name <- paste("Occurence/OCCURENCE_JULIANO_")
  }

@@

    # Join both tables
    sppdist <- sppdist %>% 
      left_join(SAU_C_data,
                by = "INDEX") %>% 
      select(TaxonKey,INDEX,everything()) %>% 
      gather("Model","Value",3:6) %>%
      mutate(Value = ifelse(is.na(Value), 0, Value))# Converting NA's to ceros

@@

    sppdist <- as.data.frame(R.matlab::readMat(SppPath))
    colnames(sppdist) <- Model

@@

# The function
EstTransIndex <- function(Spp,Model,Neighbors){

  # Get model data from spps
  SppDist <- GetSppDist(Spp,Model)

    # Estimate total spp distribution within neighbours
  Trans_Spp <- SppDist %>%
    filter(INDEX %in% Neighbors$INDEX) %>% # Filter data only located within EEZs
    mutate(Value = ifelse(Value > 0, 1,0)) %>%  # For now using SAU_C as presence absence
    group_by(TaxonKey,
             INDEX
    ) %>%
    summarise(Model_Index = mean(Value,na.rm=T)) %>% 
    filter(Model_Index > 0) %>% 
    mutate(Model_Index = Model_Index*100) %>% 
    left_join(Neighbors,
              by = "INDEX")

@@

@@

system.time(
  Spp_Trans <- bind_rows(
    mclapply(
      Exploited_Species_List[1], fun =EstTransIndex, Model = "All", Neighbors = Neighbor_List
    )
  )
)

# u ser  system elapsed 
# 72.20    9.33   91.30

### Save Test

@@

>>>>

GetSppDist=function(Spp,Model,Coord){

  Distpath <- paste(Data_Path,"Distribution_Data/",sep="")
  INDEX = seq(1,259200,1)
  # INDEX <- Coordinates$INDEX

@@

  # Occurence
  if(Model == "Occ"){
    File_Name <- paste("Occurence/OCCURENCE_JULIANO_")
  }

@@

    # Join both tables
    sppdist <- sppdist %>% 
      left_join(SAU_C_data,
                by = "INDEX") %>% 
      select(TaxonKey,INDEX,everything()) %>% 
      left_join(CoorG,
                by = "INDEX")

    # Fix coordinate system incompatibility between Gab and DBEM
    sppdist <- suppressWarnings(sppdist[order(sppdist$Latitude, rev(sppdist$Longitude),decreasing=TRUE), ] %>% 
      mutate(INDEX = seq(1,259200,1)) %>% 
      gather("Model","Value",3:6) %>%
      mutate(Value = ifelse(is.na(Value), 0, Value)) # Converting NA's to ceros
    )

@@

    sppdist <- as.data.frame(R.matlab::readMat(SppPath)) %>% 
      mutate(INDEX = INDEX,
             Species = Spp) %>% 
      left_join(Coord)

    # Fix coordinate system incompatibility between Gab and DBEM
    sppdist <- sppdist[order(sppdist$Latitude, rev(sppdist$Longitude),decreasing=TRUE), ] %>% 
      mutate(INDEX = seq(1,259200,1))

@@

#### For testing ###
# Spp <- 600004
# Model <- "All"
# Neighbors <- Neighbor_List
# Coord = CoorG
# __________________________________

# The function
EstTransIndex <- function(Spp,Model,Neighbors,Coord){

  # Get model data from spps
  SppDist <- GetSppDist(Spp,Model,Coord)

  # Estimate total spp distribution within neighbours

  Trans_Spp <- SppDist %>%
    filter(INDEX %in% Neighbors$INDEX) %>% # Filter data only located within EEZs
    mutate(Value = ifelse(Value > 0, 1,0)) %>%  # For now using SAU_C as presence absence
    group_by(TaxonKey,
             INDEX
    ) %>%
    summarise(Model_Index = mean(Value,na.rm=T)
              ) %>%
    filter(Model_Index > 0) %>%
    mutate(Model_Index = Model_Index*100) %>%
    left_join(Neighbors,
              by = "INDEX")

@@

# Gabriel's coordinate system
CoorG <- read.csv(paste(Data_Path,"Spatial_Data/coordinates_gab.csv",sep="")) %>% 
  mutate(INDEX = seq(1,259200,1))

@@

system.time(
  Spp_Trans <- bind_rows(
    mclapply(
      Exploited_Species_List[1:3], FUN =EstTransIndex, Model = "All", Neighbors = Neighbor_List,Coord = CoorG
    )
  )
)

   # user  system elapsed (For three spp in Hall1000)
 # 13.430   5.126  35.059 
### Save Test

@@

# Details

## First world Anchovies

In a first run the model suggests anchovies "Engraulis ringens" (600004) is transboundary between US-Mex-Can. I'm exploring this since all four models agree...

Found the probelm, Gab's coordinate system, hecne the data, are in different system than DBEM and SAU data...

```{r Trans_Spp_Function, eval= F ,echo = T,warning = F,message = F}

### DATA EXTRA NEEDED

# DBEM Coordinate system
Coor <- fread(paste(Data_Path,"Spatial_Data/Lon_Lat_DBEM.txt",sep=""),header = FALSE)
colnames(Coor) <- c("INDEX","Longitude","Latitude")
EEZIDs_List <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/Updated_EEZList_17June2016.xlsx",sep=""))
EEZ_CellID <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/EEZ_CellID.xlsx",sep=""))
colnames(EEZ_CellID) <- c("EEZID","INDEX")

CoorG <- read.csv(paste(Data_Path,"Spatial_Data/coordinates_gab.csv",sep="")) %>% 
  mutate(INDEX = seq(1,259200,1))

# colnames(CoorG) <- c("INDEX","Longitude","Latitude")
coords <- CoorG[order(CoorG$Latitude, rev(CoorG$Longitude),decreasing=TRUE), ]

# Explore model step by step

# Get Catch Data
Anchoveta_Dist <- GetSppDist(600004,"All",CoorG)

Anchoveta_Dist_Coor <- Anchoveta_Dist %>% 
  left_join(CoorG)

# Swap coordinate system to match DBEM and SAU
Anchoveta_Dist_Coor_Fix <- Anchoveta_Dist_Coor[order(Anchoveta_Dist_Coor$Latitude, rev(Anchoveta_Dist_Coor$Longitude),decreasing=TRUE), ]

x <- Anchoveta_Dist %>% 
  left_join(EEZ_CellID) %>% 
  left_join(EEZIDs_List) %>% 
  filter(
    # Model == "Occ",
    Value >0)

# head(x)
# Exploring using Vicky's data

ggplot(x) +
  geom_tile(aes(
    x = Longitude,
    y = Latitude,
    fill = Name,
    color = Name
  )
  ) +
  facet_wrap(~Model)

Spp_Trans <- bind_rows(
    mclapply(
      600004, FUN =EstTransIndex, Model = "All", Neighbors = Neighbor_List, Coord = CoorG
    )
  )

  # Estimate total spp distribution within neighbours

  Trans_Spp <- SppDist %>%
    filter(INDEX %in% Neighbors$INDEX) %>% # Filter data only located within EEZs
    mutate(Value = ifelse(Value > 0, 1,0)) %>%  # For now using SAU_C as presence absence
    group_by(TaxonKey,
             INDEX
    ) %>%
    summarise(Model_Index = mean(Value,na.rm=T)) %>%
    filter(Model_Index > 0) %>%
    mutate(Model_Index = Model_Index*100) %>%
    left_join(Neighbors,
              by = "INDEX")

    Test <- Trans_Spp %>% 
      filter(Name == "USA (West Coast)" & Model_Index ==100)

 # Explore only SAU catch data

# Get Catch Data
Anchoveta_Dist <- GetSppDist(Spp,"SAU_C")

# Get Mex, Can and USA W coast Grid
Anchoveta_Dist_EEZ <- Anchoveta_Dist %>%

Anchoveta_NA <-  Anchoveta_Dist%>% 
  filter(INDEX %in% North_America$INDEX) %>% 
  gather("Year","Tonnes",CATCH.1:CATCH.65) %>% 
  group_by(Year) %>% 
  summarise(Sum_T = sum(Tonnes, na.rm =T))
